---
title: "p8105_hw3_eih2108"
author: "Ekaterina Hofrenning"
date: '2023-10-10'
output: github_document
---

```{r}
library(readr)
library(tidyverse)
library(scrutiny)
library(knitr)
```

# Question 1

```{r}
# install
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data(nyc_airbnb)


nyc_airbnb <- nyc_airbnb %>%
  rename(borough = neighbourhood_group) %>%
  mutate(stars = review_scores_location / 2)

nyc_airbnb %>%
  group_by(host_id, borough) %>%
  summarize(num = n()) %>%
  arrange(desc(num)) %>%
  head(10)
```


# Question 2

```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data("brfss_smart2010")
```

First, do some data cleaning:
- format the data to use appropriate variable names
- focus on the “Overall Health” topic: include only responses from “Excellent” to “Poor” and organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r, warning = F}
brfss <-
brfss_smart2010 %>% 
  separate(Locationdesc, into = c("State", "County"), sep = "-") %>% # create appropriate state & cty vars
  select(-c(Locationabbr)) %>%
  mutate(overall_health = as.factor(case_when(Topic == "Overall Health" ~ Response))) %>%
  mutate(Response = as.factor(Response))
```

Using this dataset, do or answer the following (commenting on the results of each):

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss %>%
  filter(Year == 2002) %>%
  group_by(State) %>%
  summarize(num = n_distinct(GeoLocation)) %>%
  filter(num > 7)

brfss %>%
  filter(Year == 2010) %>%
  group_by(State) %>%
  summarize(num = n_distinct(GeoLocation)) %>%
  filter(num > 7)
```

Construct a dataset that is limited to Excellent responses, and contains year, state, and a variable that averages the data_value across locations within a state. 

```{r, warning = F, message = F}
brfss_avg <- brfss %>%
    filter(overall_health == "Excellent") %>% # limit to Excellent
    select(Year, State, GeoLocation, Data_value) %>%
    group_by(Year, State) %>%
    summarise(mean_data_value = mean(Data_value)) %>% # average data_value across geo-locations within a state
    ungroup()
```

Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r, warning=FALSE, message=FALSE}
ggplot(brfss_avg, aes(x = Year, y = mean_data_value, group = State)) +
  geom_point(size = 1) + 
  geom_line() +
  labs(y = "Average Value")
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss %>%
  filter(Year %in% c(2006, 2010), 
         State == "NY ",
         !is.na(overall_health)) %>%
  select(Year, County, overall_health, Data_value) %>%
  arrange(Year, County, overall_health) %>%
  ggplot(aes(x = overall_health, y = Data_value)) +
  geom_boxplot() +
  facet_wrap(~Year) +
  labs(x = "Overall Health", y = "Data value")
```

# Question 3

Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r, warning = F, message=FALSE}
covar <- read_csv("./data/nhanes_covar.csv") %>%
  row_to_colnames(row = 4) %>%
  slice_tail(n = 250) %>%
  mutate(SEQN = as.integer(SEQN), 
         sex = as.factor(case_when(sex == 1 ~ "Male", sex == 2 ~ "Female")), 
         age = as.integer(age), 
         BMI = as.double(BMI), 
         education = as.factor(case_when(education == 1 ~ "Less than high school",
                                         education == 2 ~ "High school equivalent",
                                         education == 3 ~ "More than high school"))) %>%
  na.omit() %>%
  filter(age >= 21)

accel <- read_csv("./data/nhanes_accel.csv")

nhanes <- left_join(covar, accel, by = "SEQN") # merge together
```

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

The table below describes the education category and sex breakdowns in the data. The highest education category and sex combination is Females with more than a high school degree (59 people). The lowest education and sex combination is Females with a high school or high school-equivalent degree.

As seen in the graphic below, females completing high school (or high school equivalent) have a higher median age than males of the same education level in our dataset. Similarly, females with less than a high school degree have a slightly higher median age than males of the same education level in our dataset. Males with more than a high school degree have a higher median age than females with the same education level.

```{r, warning=FALSE, message=FALSE}
# create table
nhanes %>% 
  group_by(education, sex) %>%
  summarize(`Number of People` = n()) %>%
  ungroup() %>%
  rename(Education = education, Sex = sex) %>%
  kable(digits = 1)

# create boxplots (sex, age, education)
nhanes %>%
  select(SEQN, sex, age, education) %>%
  ggplot(aes(x = sex, y = age)) +
  geom_boxplot() +
  facet_wrap(~education)
```

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r, message=FALSE}
nhanes_sum <- nhanes %>%
  mutate(activity_per_day = rowSums(select(., contains("min")))) %>%
  select(SEQN:education, activity_per_day)

ggplot(nhanes_sum, aes(x = age, y = activity_per_day, color = sex)) +
  geom_point() +
  geom_smooth(se = F) +
  facet_wrap(~education) +
  labs(y = "Activity level per day", x = "Age")
```

Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}
nhanes %>% 
  pivot_longer(cols = contains("min"), names_to = "minute", values_to = "activity") %>%
  separate(col = minute, "n", into = c("min_char", "min")) %>%
  mutate(min = as.integer(min)) %>%
  ggplot(aes(x = min, y = activity, color = sex)) +
  geom_point(alpha = .1) +
  geom_smooth(se = F) +
  facet_wrap(~education) +
  labs(x = "Minutes", y = "Activity level")
```



```{r}
nhanes %>% 
  pivot_longer(cols = contains("min"), names_to = "minute", values_to = "activity") %>%
  separate(col = minute, "n", into = c("min_char", "min")) %>%
  mutate(min = as.integer(min)) %>%
  ggplot(aes(x = min, y = activity, color = sex)) +
  geom_point(alpha = .1) +
  geom_smooth(se = F) +
  facet_wrap(~education)
```

