---
title: "p8105_hw3_eih2108"
author: "Ekaterina Hofrenning"
date: '2023-10-10'
output: html_document
---



```{r}
# install
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data(nyc_airbnb)
library(tidyverse)

nyc_airbnb <- nyc_airbnb %>%
  rename(borough = neighbourhood_group) %>%
  mutate(stars = review_scores_location / 2)

nyc_airbnb %>%
  group_by(host_id, borough) %>%
  summarize(num = n()) %>%
  arrange(desc(num)) %>%
  head(10)
```


# Question 2

```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data("brfss_smart2010")
```

First, do some data cleaning:
- format the data to use appropriate variable names
- focus on the “Overall Health” topic: include only responses from “Excellent” to “Poor” and organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
m1 <-
brfss_smart2010 %>% 
  separate(Locationdesc, into = c("State", "County"), sep = "-") %>% # create appropriate state and county vars
  select(-c(Locationabbr)) %>%
  mutate(overall_health = as.factor(case_when(Topic == "Overall Health" ~ Response))) %>%
  select(Year, State, County, Topic, Response, overall_health)
```

Using this dataset, do or answer the following (commenting on the results of each):

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
m1 %>%
  filter(Year == 2002) %>%
  group_by(State) %>%
  summarize(num = n_distinct(GeoLocation)) %>%
  filter(num > 7)

m1 %>%
  filter(Year == 2010) %>%
  group_by(State) %>%
  summarize(num = n_distinct(GeoLocation)) %>%
  filter(num > 7)
```

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

```{r}
m1 %>%
  
```


Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.



```{r}
num_simulations <- 100000
rewards <- numeric(num_simulations)

# Simulate
for (i in 1:num_simulations) {
  reward <- 2^sample(1:100, 1, replace = TRUE)  # Simulate up to 100 flips
  rewards[i] <- reward
}

mean_reward <- mean(rewards)
cat("Mean Reward:", mean_reward, "\n")
```




# Question 3
